---
title: "Analysis of Targeted Example Presentation"
author: "Kenneth C. Arnold"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("readr")
library("tidyverse")
library("ggbeeswarm")
library("afex")
afex::set_sum_contrasts()
library("ARTool")
library("emmeans")
```

```{r read, include=FALSE}
BATCH <- "idea3r5"
DATA_PATH <- paste0("../data/analyzed/idea/", BATCH, "/")
do_read_csv <- function(filename, ...) {
  data <- read_csv(paste0(DATA_PATH, filename), ...);
  stop_for_problems(data);
  data;
}

block_factor <- col_factor(c("0", "1", "2"));
all_trialLevel <- do_read_csv("trial.csv",col_types = cols(block = block_factor));
all_trialLevel$condition <- factor(all_trialLevel$condition)
trialLevel <- all_trialLevel %>% subset(should_exclude==FALSE);

all_expLevel <- do_read_csv("experiment.csv")
expLevel <- all_expLevel %>% subset(any_trial_excluded==FALSE);
all_ideaLevel <- do_read_csv("ideas.csv", col_types=cols(block=block_factor));
ideaLevel <- all_ideaLevel %>% subset(should_exclude==FALSE);
```

# Summary

```{r counts}
all_expLevel %>% group_by(any_trial_excluded) %>% summarise(n=n_distinct(participant), total_time=median(total_time))
```

```{r time_taken_graph}
ggplot(all_expLevel, aes(x=total_time)) + geom_histogram(bins=30) + facet_wrap(vars(any_trial_excluded), nrow = 2)
```

```{r checkRandomization}
#data %>% group_by(block, condition) %>% summarise(n=n_distinct(participant))
all_expLevel %>% group_by(any_trial_excluded, condition_order) %>% summarize(n=n())
```

TODO: Why was randomSents-first so heavily excluded?

# Did the system basically work?

- Inspirations got shown, on trials other than norecs.


```{r systemFailures}
all_trialLevel %>% group_by(condition, any_requests_returned_empty) %>% summarise(n())
```


TODO: Did people use some word that was a cue word while it was visible?

## How many times did people request inspiration?

```{r numRequests}
trialLevel %>% group_by(condition) %>% summarise_at(c("num_inspiration_requests"), mean)
```

```{r requests}
trialLevel %>% 
  ggplot(aes(x=condition, y=num_inspiration_requests)) +
    geom_beeswarm() +
    facet_wrap("block")
```


## Fluency

How many ideas did they have after 5 minutes?

```{r fluency}
labeledBar <- list(
  stat_summary(geom="bar", fun.y=mean, position="dodge"),
    stat_summary(geom="errorbar", fun.data=mean_se, position=position_dodge(width=.9), width=.4),
    scale_y_continuous(expand=expand_scale(mult=c(0,.05))), # remove the extra spacing at 0
    stat_summary(geom="text", fun.y=mean,
                 aes(label=sprintf("%1.2f", ..y..)),
                 position=position_dodge(width=.9),
                 hjust="middle", vjust="bottom"))

trialLevel %>% ggplot(aes(x=condition, y=fluency_5)) + labeledBar# + facet_wrap("block")
```


Once we remove obvious outliers, the fluency results are very similar across conditions.
```{r timeTillN}
ideaLevel$outlier <- ideaLevel$since_start > quantile(ideaLevel$since_start, .75) + 1.5 * IQR(ideaLevel$since_start);


ideas_by_time <- ideaLevel %>% group_by(participant, block) %>% mutate(nth=row_number()) %>% subset(nth <= 20);
ideas_by_time %>% 
  group_by(nth) %>% filter((since_start < quantile(since_start, .75) + 1.5*IQR(since_start))) %>% 
  ggplot(aes(x=nth, y=since_start, color=condition)) +
    stat_smooth() + geom_beeswarm() + 
    facet_wrap(vars(block)) +
    #coord_flip() +
    labs(x="Num questions generated", y="Seconds since start")
```

```{r timeTillNAnalysis, include=F}
# Not including this since it's unclear how to do it correctly.
#mixed(nth ~ since_start * condition * block + (block | participant), data=ideas_by_time)
```

TODO: Fluency measures from past papers.
TODO: Total time needed to write 120 words of final text.



## Subjective Experience

```{r include=F}
multiPlot <- function(measures.to.vis, geom=labeledBar) {
  trialLevel %>%
    select(participant, condition, block, !!!measures.to.vis) %>%
    gather(measure, value, !!!measures.to.vis) %>%
    mutate(measure=factor(measure, levels=measures.to.vis)) %>%
    ggplot(aes(x=condition, y=value)) +
      geom +
      facet_wrap(vars(measure, block), ncol=3, scales="free", strip.position="top") +
      labs(x="condition", y="") +
      theme(
        strip.background = element_blank(),
        panel.spacing.y = unit(2, "lines"),
        axis.line.x = element_blank(),
        legend.position = "bottom")
}
```


```{r helpedOverall, fig.cap="The system was helpful overall."}
multiPlot(c("brainstorm_system_helped"), geom=geom_boxplot())
```

```{r usePaper, fig.cap="I would have done better using paper."}
multiPlot(c("brainstorm_distracting"))#, geom=geom_boxplot())
```


```{r process, fig.cap="I felt like I could come up with ideas easily. // I sometimes felt stuck."}
multiPlot(c(
    "brainstorm_fluent",
    "brainstorm_stuck"
));
```

```{r system, fig.cap="The inspirations gave me new ideas. // The inspirations were relevant. // I used the inspirations that were given."}
multiPlot(c(
    "brainstorm_sysGaveIdeas",
    "brainstorm_sysRelevant",
    "brainstorm_usedInspirations"
  ));
```


```{r subjectiveOverall}
trialLevel <- trialLevel %>% 
  mutate(brainstorm_overall=brainstorm_system_helped+brainstorm_sysRelevant+brainstorm_sysGaveIdeas+brainstorm_usedInspirations - brainstorm_distracting)

trialLevel %>% 
  ggplot(aes(x=condition, y=brainstorm_overall)) + labeledBar + facet_wrap(vars(block))
```


Targeted recommendations are better than nothing but not better than random.
```{r relevance}

ana <- function(formula, use_art=T, data=NA) {
  if (is.na(data)) data <- trialLevel;
  res <- list()
  if(use_art) {
    m <- art(formula, data=data)
    res$alignSummary <- summary(m)
    res$m <- anova(m)
    res$emm <- emmeans(artlm(m, "condition"), "condition", contr="pairwise", adjust="holm")
    resids <- residuals(m)
  } else {
    res$m <- mixed(formula, data=data, method="KR")
    res$emm <- emmeans(res$m, "condition", by=by, contr="pairwise", adjust="holm")
    resids <- residuals(res$m$full_model)
  }
  res$shapiroResid <- shapiro.test(resids)
  car::qqPlot(resids, groups=data$condition)
  res
}
list(
  relevant=ana(brainstorm_sysRelevant ~ condition * block + (1|participant)),
  gaveIdeas=ana(brainstorm_sysGaveIdeas ~ condition * block + (1|participant)),
  usedInspirations=ana(brainstorm_usedInspirations ~ condition * block + (1|participant)))
```
```{r analyzeSubjectiveOverall}
ana(brainstorm_overall ~ condition * block + (1|participant))
```

```{r pickExamples}
trialLevel %>% subset(condition=="highlightedSents") %>% select(participant, brainstorm_overall) %>% arrange(brainstorm_overall)# %>% pull(participant)
```

# Relevance

How relevant were the generated ideas to the writing task? TODO. 

# Breadth of Ideas Generated

We use word-vector-based metrics to compute the breadth of ideas generated. (This is in a Jupyter Notebook. And no difference between conditions.)


# Subjective Experience


  - ... of the system: did writers find the system's suggestions relevant? useful? giving ideas?
  - ... of the task: did they feel stuck?

## Comments

```{r comments}
trialLevel %>% group_by(condition) %>% summarize(comments=paste(brainstorm_whenRequest, collapse = " // ")) %>% knitr::kable()
```

## Self-Efficacy

```{r computeSelfEfficacy}
for (block in c(0:2)) {
  for (aspect in c("WhatToSay", "HowToSay")) {
    expLevel[paste0("pre2mid_", aspect, '_', block)] <- 
      expLevel[paste0("mid-selfEfficacy", aspect, "-", block)] - expLevel[paste0("pre-selfEfficacy", aspect, "-", block)];
  }
}
selfEfficacyPre2Mid <- expLevel %>% 
  select(participant, starts_with("pre2mid")) %>% 
  gather(sedelta, delta, -participant) %>% 
  extract(sedelta, c("aspect", "block"), "pre2mid_(\\w+)_(\\d)", convert=T)
selfEfficacyPre2Mid <- selfEfficacyPre2Mid %>% mutate(block=factor(block))
```

```{r plotSelfEfficacy}
seDeltaByTrial <- trialLevel %>% select(participant, block, condition) %>% 
  left_join(selfEfficacyPre2Mid, by=c("participant", "block"))
seDeltaByTrial %>%
  filter(!is.na(delta)) %>% 
  ggplot(aes(x=condition, y=delta)) +
    labeledBar+
    #geom_beeswarm() +
    facet_wrap(vars(aspect, block))
```

# Final Writing

## Informativeness

```{r nll}
multiPlot(c("nll_total_lower", "nll_per_char"))
```

```{r nll_simple}
trialLevel %>% ggplot(aes(x=condition, y=nll_per_char)) + geom_boxplot() + facet_wrap(vars(block))
```

```{r nll_analysis}
mixed(nll_per_char ~ condition * block + (1|participant), data=trialLevel)
```

## Quality / Style



# Appendix

For more details on using R Markdown see <http://rmarkdown.rstudio.com>.
